\documentclass[10pt, a4paper]{article}
\clubpenalty1000000
\widowpenalty1000000
\usepackage{lrec2006}
\usepackage{graphicx}
\usepackage[pdfborder={0 0 0}]{hyperref}
\usepackage{listings}
\usepackage{amsmath,amssymb,colortbl,xcolor}
\usepackage{tipa}

\title{An Interactive Visualization of Cross-Linguistic Colexification Patterns}

%\name{Thomas Mayer$^{1}$, Johann-Mattis List$^{1}$, Anselm Terhalle$^{2}$}
%
%\address{ 
%               $^{1}$Philipps University of Marburg, $^{2}$Heinrich-Heine University of D\"usseldorf \\
%               $^{1}$Deutschhausstra{\ss}e 3, 35037 Marburg, $^{2}$Universit\"atsstra{\ss}e 1, 40225 D\"usseldorf \\
%               thomas.mayer@uni-marburg.de, mattis.list@uni-marburg.de, terhalle@phil.uni-duesseldorf.de\\}


\abstract{
In this paper, we present an interactive web-based visualization for the CLiCS database, an online
resource for synchronic lexical associations (\emph{colexification patterns})
in over 200 language varieties. The associations cover
1,288 concepts and represent the tendency for concepts to be expressed by the same words in
the same languages and language varieties of the world. 
The complexity of the network structure in the CLiCS database calls for
a visualization component that makes it easier for researchers to explore the patterns of
cross-linguistic colexifications. The network is represented as a force-directed graph and features a
number of interactive components that allow the user to get an overview of the overall structure
while at the same time providing the opportunity to look into the data in more detail. An integral
part of the visualization is an interactive listing of all languages that contribute to the strength
of a given pattern of colexification. Each language in the list is thereby attributed a different
color depending on its genealogical or areal affiliation. In this way, given associations can be
inspected for genealogical or areal bias. 
\\ \newline \Keywords{Interactive visualization, colexification, cross-linguistic database}}




\begin{document}

\maketitleabstract

\section{Introduction}
% stress the importance of the visualization as added value to the resource; what can be seen with the visualization that would be otherwise hard to notice

What does ``good'' have to do with ``beautiful''? Logically, not everything that is good is beautiful, and not everything that is beautiful is good. However, people seem to associate these concepts quite strongly, as linguistic data suggest: ``good'' and ``beautiful'' are expressed by identical word forms in 27 languages from 8 different language families. To assess the cognitive, linguistic, and cultural implications of this fact correctly, additional information would be useful: Where are these languages located on the globe? How are they distributed amoung the 8 families? Which other concepts are verbalized by the same form as ``good'' and ``beautiful''?

Synchronic lexical association or \textsl{colexification}, i.e.~the verbalization of two or more concepts by means of the same form in a given language, is an important source of information for investigations in cognitive linguistics, linguistic typology, and historical semantics. In this paper, we present an interactive web-based visualization for CLiCs, an online database that contains a large crosslinguistic data set on colexifications worldwide. We provide a brief outline of CLiCs, describe the functionalities of the visualization, and show that the visualization is an indispensible tool that enables researches to get an overview of the data and concisely plan further quantitative analyses according to their needs.

\section{CLiCS}

\input{sections/clics}

\section{Visualization}
% Thomas

\input{sections/vis}

\section{Conclusions and future work}

The size and complexity of today's language resources call for a data preparation pipeline that enables researchers to find meaningful patterns among the multitude of different factors that can be taken into consideration. In our view, such a data preparation pipeline necessarily consists of two major parts, both of which are illustrated in the present paper. On the one hand, methods and techniques from data mining or computational linguistics help to detect basic trend or groups of similar objects in the search space. On the other hand, the resulting groups or trends are mapped to visual variables in order to make interesting observations readily accessible to human perception. 

The CLiCs database contains a wealth of information about colexification patterns in the languages of the world. Manually inspecting the large amount of connections in the database, however, is a laborious and time-consuming task that allows for a detailed exploration of individual links but does not capture overall trends in the data. This paper presents an attempt to combine the advantages of human inspection with the strength of a computational approach \cite{Keim2008}. 

The CLiCs visualization features an automatic preprocessing of the colexification links into so-called communities of the graph, groups of highly connected nodes that reveal a meaningful overall trend in the worldwide patterns of lexical associations. The communities are then graphically represented in a force-directed graph that shows all connections within the various concepts that are included. Interactive components in the visualization allow for a more detailed view of associations on the level of the languages that contribute to the colexification. 
Mapping the genealogical and areal information on individual languages to colors enables an at-a-glance evaluation of potentially interesting trends in individual colexifications (see the case study in Section \ref{case study}). 
In this way, users can get an overview of the general trends in the data and at the same time have the opportunity to directly inspect the lexical associations. 

In future work, we plan to enhance the visualization tool with further interactive components that allow for a better overview of the complete network of colexifications (shown in Figure \ref{fig:clics_full}) facilitate the detection of genealogical or areal trends in the database. The idea is to integrate a sunburst visualization \cite{Sunburst}  for the genealogical information in order to enable a better overview of the language families that are involved in a given colexification pattern.\footnote{See \cite{MayerLanguageExplorer} for an example of using sunburst displays to represent the hierarchical structure of language families.} In addition, we intend to equip the user interface with further interactive components that allow users to explore the database from different perspectives (e.g., compare individual languages in terms of shared lexical associations). All components will be made publicly available online for the (linguistic) research community. 


\bibliographystyle{lrec2006}
\bibliography{clicsvis}

\end{document}

